Notes on the hydrogen-bond auto-correlation

### 2014-10-16 writedown/auto.correlations/h.bond.notes/main.tex
fix bugs

\begin{abstract}
  In this note I describe the implmentational details of the hydrogen-bond auto-correlations.
\end{abstract}

@@

\section{Definitions}
Here I use the definition of hydrogen-bond population from
Ref.~\cite{luzar1996effect, luzar1996hydrogen}.  The population $h(t)$
is defined to be 1 if a certain pair of molecules forms hydrogen-bond,
otherwise is defined to be 0. Therefore, the averaged total number of
hydrogen-bond in an equilibrium system is $\frac12 N(N-1)\langle
h\rangle$, where $\langle\cdot\rangle$ denotes the ensemble average.
The auto-correlation
\begin{align}
  c(t) = \frac{\langle h(0) h(t)\rangle}{\langle h\rangle}
\end{align}
is studied. The physical meaning is that given a hydrogen-bond formed
at time $0$, the probability of observing a hydrogen-bond between the
same pair at time $t$.
The rate of relaxation to equilibrium is of paricular interest:
\begin{align}\label{eqn:k-def}
  k(t) &= -\frac{dc}{dt}\\\label{eqn:k-exp}
  &= -\frac{\langle \dot h(0) [1 - h(t)] \rangle}{\langle h\rangle}
\end{align}
where $\dot h(0) = (dh/dt)\vert_{t=0}$. The Eq.~\eqref{eqn:k-exp} holds because
\begin{align} \label{eqn:tmp3}
  \langle h(0) \dot h(t) \rangle &= - \langle \dot h(0) h(t)\rangle \\\label{eqn:tmp4}
  \langle \dot h(0) \rangle &= 0
\end{align}
Eq.~\eqref{eqn:tmp4} is obvious because the system is in equilibrium. Eq.~\eqref{eqn:tmp3} holds because
\begin{align}
  0 = \frac{d [ h(\tau) h(\tau+t) ] }{d\tau} =  \langle h(\tau) \dot h(\tau+t)\rangle  + \langle \dot h(\tau) h(\tau+t)\rangle.
\end{align}
Letting $\tau \rightarrow 0$ on both sides yields Eq.~\eqref{eqn:tmp3}.

@@

The rate of relaxation to equilibrium $k(t)$ can be calculated by
definition~\eqref{eqn:k-def}, or calculated from Eq.~\eqref{eqn:k-exp}
with an explicit estimation of the function $\dot h$. We want to discuss the second way.

@@

the O--H...O angle is less than a cut-off $\theta_c$. To mathematically define the function $h$, we introduce

@@

If we denote the DOFs of two water molecules by $\vect x_1$ and $\vect x_2$, then the hydrogen-bond population function $h$ is expressed by

@@

the O--H...O angle. In practice, it is not easy to calcualte the
angle, but the cosine angle is relatively easy. Therefore we use the

@@

If we notice the fact that the DOFs of the molecules are functions of

@@

We compute the two derivatives in the equation:

@@

      &\frac 1{2\Delta L}, &\qquad& -\frac{\Delta L}{2} \leq x < \frac{\Delta L}{2}, \\

@@

\vect x_2)/dt$, we fristly derive the derivative w.r.t.~the distance
between any two positions $\vert \vect r_1 - \vect r_2\vert$:

@@

  \frac{d\vert \vect r_1(t) - \vect r_2(t)\vert}{dt}

@@

  \frac{( \vect r_1(t) - \vect r_2(t) )}{  \vert \vect r_1(t) - \vect r_2(t)\vert} \cdot
  \frac{ d( \vect x_1(t) - \vect x_2(t) )}{dt}  

@@

  \frac{( \vect r_1(t) - \vect r_2(t) ) \cdot (\vect v_1(t) - \vect v_2(t))}{  \vert \vect r_1(t) - \vect r_2(t)\vert} 

@@

where $\vect r_{12}^{\oxy\oxy}$ denotes the oxygen--oxygen distance between molecules 1 and 2. For the derivative of the cosine angle:

@@

>>>>

\begin{abstract}
  In this note I describe the implementation details of the hydrogen-bond auto-correlations.
\end{abstract}

@@

\section{Definitions}
Here I use the definition of hydrogen-bond population from
Ref.~\cite{luzar1996effect, luzar1996hydrogen}.  The population $h(t)$
is defined to be 1 if a certain pair of molecules forms hydrogen-bond,
otherwise is defined to be 0. Therefore, the averaged total number of
hydrogen-bond in an equilibrium system is $\frac12 N(N-1)\langle
h\rangle$, where $\langle\cdot\rangle$ denotes the ensemble average.
The auto-correlation
\begin{align}
  c(t) = \frac{\langle h(0) h(t)\rangle}{\langle h\rangle}
\end{align}
is studied. The physical meaning is that given a hydrogen-bond formed
at time $0$, the probability of observing a hydrogen-bond between the
same pair at time $t$.
The rate of relaxation to equilibrium is of particular interest:
\begin{align}\label{eqn:k-def}
  k(t) &= -\frac{dc}{dt}\\\label{eqn:k-exp}
  &= -\frac{\langle \dot h(0) [1 - h(t)] \rangle}{\langle h\rangle}
\end{align}
where $\dot h(0) = (dh/dt)\vert_{t=0}$. The Eq.~\eqref{eqn:k-exp} holds because
\begin{align} \label{eqn:tmp3}
  \langle h(0) \dot h(t) \rangle &= - \langle \dot h(0) h(t)\rangle \\\label{eqn:tmp4}
  \langle \dot h(0) \rangle &= 0
\end{align}
Eq.~\eqref{eqn:tmp4} is obvious because the system is in equilibrium. Eq.~\eqref{eqn:tmp3} holds because
\begin{align}
  0 = \frac{d [ h(\tau) h(\tau+t) ] }{d\tau} =  \langle h(\tau) \dot h(\tau+t)\rangle  + \langle \dot h(\tau) h(\tau+t)\rangle.
\end{align}
Letting $\tau \rightarrow 0$ on both sides yields Eq.~\eqref{eqn:tmp3}.

@@

The rate of relaxation to equilibrium $k(t)$ can be calculated by
definition~\eqref{eqn:k-def}, or calculated from Eq.~\eqref{eqn:k-exp}
with an explicit estimation of the function $\dot h$. I want to discuss the second way.

@@

the O--H...O angle is less than a cut-off $\theta_c$. To mathematically define the function $h$, I introduce

@@

If I denote the DOFs of two water molecules by $\vect x_1$ and $\vect x_2$, then the hydrogen-bond population function $h$ is expressed by

@@

the O--H...O angle. In practice, it is not easy to calculate the
angle, but the cosine angle is relatively easy. Therefore I use the

@@

If noticing the fact that the DOFs of the molecules are functions of

@@

The two derivatives in the equation is computed by:

@@

      &\frac 1{\Delta L}, &\qquad& -\frac{\Delta L}{2} \leq x < \frac{\Delta L}{2}, \\

@@

\vect x_2)/dt$, I firstly derive the time-derivative of the distance
between any two positions $r_{12} = \vert \vect r_{12}\vert  = \vert \vect r_2 - \vect r_1\vert$:

@@

  \frac{d\vert \vect r_2(t) - \vect r_1(t)\vert}{dt}

@@

  \frac{( \vect r_2(t) - \vect r_1(t) )}{  \vert \vect r_2(t) - \vect r_1(t)\vert} \cdot
  \frac{ d( \vect x_2(t) - \vect x_1(t) )}{dt}  

@@

  \frac{( \vect r_2(t) - \vect r_1(t) ) \cdot (\vect v_2(t) - \vect v_1(t))}{  \vert \vect r_2(t) - \vect r_1(t)\vert} 

@@

where $r_{12}^{\oxy\oxy}$ denotes the oxygen--oxygen distance between molecules 1 and 2. For the derivative of the cosine angle:

@@

The meaning of the notations should be self-explanatory.
